Ride at Your Own Risk: Insights into Toronto’s Bicycle Theft Hotspots

### 2024-11-11 paper/paper.Rmd
add background image

---
title: "Ride at Your Own Risk: Insights into Toronto’s Bicycle Theft Hotspots"
subtitle: "[To Be Updated....]"
author: "Yingke He, Ziheng Zhong"
thanks: "Code and data are available at: [https://github.com/iJustinn/Toronto_Bicycle_Thefts/tree/main](https://github.com/iJustinn/Toronto_Bicycle_Thefts/tree/main)."
date: today
output: html_document
bibliography: references.bib
---

<style>
body {
  background-color: #f0f8ff; /* AliceBlue color */
}
</style>

@@

# you are welcome to use either the CSV file or the GeoJSON file
# GeoJSON format maybe useful if you would like to combine
# an external geospatial dataset
bike_thefts_csv <- read_csv(here("data/clean_data.csv")) |>
  mutate(
    occurence_date = as_date(
      OCC_DATE, format = "%a, %d %b %Y %H:%M:%S GMT"
    ),
    report_date = as_date(
      REPORT_DATE, format = "%a, %d %b %Y %H:%M:%S GMT"
    )
  )
bike_thefts_sf <- read_sf(here("data/Bicycle_Thefts_Open_Data.geojson")) |>
  mutate(
    occurence_date = as_date(
      OCC_DATE, format = "%a, %d %b %Y %H:%M:%S GMT"
    ),
    report_date = as_date(
      REPORT_DATE, format = "%a, %d %b %Y %H:%M:%S GMT"
    )
  )
```
Map
```{r, include = TRUE, warning = FALSE, message = FALSE}
# Load cleaned coordinates data
coordinates_data <- read.csv(here("data/cleaned_coordinates_data.csv"))

# Plot the map with bike theft locations data
ggplot() +
  geom_sf(data = streets, color = "gray", size = 0.3, alpha = 0.6) + # OSM street map layer
  geom_point(
    data = coordinates_data, aes(x = longitude, y = latitude),
    size = 0.5, color = "red"
  ) + # Bike theft points
  labs(
    title = "Map of Bicycle Theft Locations in Toronto",
    x = "Longitude", y = "Latitude"
  ) +
  coord_sf(
    xlim = c(-79.6303856025883, -79.1180336515019),
    ylim = c(43.5822069220627, 43.8554571861712), expand = FALSE
  ) +
  theme_minimal() +
  theme(
    panel.background = element_blank(), # Remove panel background
    panel.grid.major = element_blank(), # Remove major grid lines
    panel.grid.minor = element_blank(), # Remove minor grid lines
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
# introduction
write an intro to convey the issue
the goal of the article :be careful of bikes

@@

@@

# Plot for Bikes Found with scaled heights
plot_found <- ggplot(data, aes(x = OCC_MONTH_NUM, y = OCC_YEAR, 
                               height = (Found / max_found) * 4, group = OCC_YEAR)) +
  geom_ridgeline(
    fill = "#05595B", color = "white", alpha = 0.5, scale = 1, linewidth = 0
  ) +
  labs(y = "", x = NULL) +  # Remove x-axis title
  theme_minimal() +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(size = 0.5),
    panel.border = element_blank(),
    plot.title = element_text(size = 10, color = "#05595B", hjust = -0.1, vjust = 1.5),  # Smaller title
    panel.background = element_rect(fill = "#f0f8ff", color = NA),  # Light blue background for plot area
    plot.background = element_rect(fill = "#f0f8ff", color = NA)    # Light blue background for overall plot
  ) +
  scale_x_discrete(labels = month.abb) +
  ggtitle("Bikes Found")  # Add title for Bikes Found

# Plot for Bikes Stolen with scaled heights and no borders
plot_stolen <- ggplot(data, aes(x = OCC_MONTH_NUM, y = OCC_YEAR, 
                                height = (Stolen / max_stolen) * 4, group = OCC_YEAR)) +
  geom_ridgeline(
    fill = "#603601", color = "white", alpha = 0.5, scale = 1, linewidth = 0
  ) +
  labs(y = "", x = NULL) +  # Remove x-axis title
  theme_minimal() +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(size = 0.5),
    panel.border = element_blank(),
    panel.background = element_rect(fill = "#f0f8ff", color = NA),  # Light blue background for plot area
    plot.background = element_rect(fill = "#f0f8ff", color = NA),   # Light blue background for overall plot
    plot.title = element_text(size = 10, color = "#603601", hjust = -0.1, vjust = 1.5)  # Smaller title
  ) +
  scale_x_discrete(labels = month.abb) +
  ggtitle("Bikes Stolen")  # Add title for Bikes Stolen

@@

```{r interactive_weekly_plot, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=7, fig.height=4}
library(plotly)
library(readr)
library(dplyr)
library(here)

# Add layout with the slider and set background colors
fig <- fig %>%
  layout(
    xaxis = list(title = ""),
    yaxis = list(title = "Number of Thefts"),
    sliders = list(
      list(
        active = 0,
        currentvalue = list(prefix = "Day: "),
        steps = steps
      )
    ),
    plot_bgcolor = "#f0f8ff",     # Set background color for the plotting area
    paper_bgcolor = "#f0f8ff"     # Set background color for the entire plot
  )

@@

```{r animated_bar_plot, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=7, fig.height=4}
# Libraries
library(ggplot2)
library(gganimate)
library(readr)
library(here)
library(dplyr)

@@

  # Set background colors for panel and plot
  theme(
    panel.background = element_rect(fill = "#f0f8ff", color = NA),
    plot.background = element_rect(fill = "#f0f8ff", color = NA)
  ) +

@@

# Create the animated plot
plot <- data %>%
  ggplot(aes(x = OCC_MONTH_NUM, y = Number_of_Thefts, group = as.factor(OCC_YEAR), color = as.factor(OCC_YEAR))) +
  geom_line() +
  geom_point() +
  scale_color_viridis_d() +
  ylab("Number of Bikes Stolen") +
  xlab("Month") +
  scale_x_continuous(breaks = 1:12, labels = c("January", "February", "March", "April", "May", "June", 
                                               "July", "August", "September", "October", "November", "December")) +
  theme_ipsum() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.background = element_rect(fill = "#f0f8ff", color = NA),
    plot.background = element_rect(fill = "#f0f8ff", color = NA)
  ) +
  transition_reveal(OCC_MONTH_NUM) +
  labs(color = "Year")

@@

>>>>

---
thanks: "Code and data are available at: [https://github.com/iJustinn/Toronto_Bicycle_Thefts/tree/main](https://github.com/iJustinn/Toronto_Bicycle_Thefts/tree/main)."
output: html_document
bibliography: references.bib
---
---
output: html_document
---

<!-- Background image -->
<style>
body {
  background-image: url('bikes1.jpg');
  background-size: cover;
  background-position: center;
  background-attachment: fixed;
  color: white;
  font-family: Arial, sans-serif;
}

.title {
  font-size: 60px;
  font-weight: bold;
  color: #FFC107; /* Yellow color */
  margin-top: 100px;
  text-align: left;
  font-family: 'Georgia'
}

.subtitle {
  font-size: 35px;
  font-weight: normal;
  color: white;
  text-align: left;
  margin-top: 10px;
  font-family: 'Open Sans', sans-serif;
}

.description {
  font-size: 18px;
  font-weight: normal;
  color: white;
  line-height: 1.6;
  text-align: left;
  max-width: 600px;
  margin-top: 20px;
  font-family: 'Lora'
}

.author-date {
  font-size: 18px;
  font-weight: normal;
  color: white;
  text-align: left;
  margin-top: 30px;
  font-style: italic;
}
</style>

<div class="title">Ride at Your Own Risk: Insights into Toronto’s Bicycle Theft Hotspots</div>

<div class="subtitle">Navigating Toronto’s streets with eyes wide open, this data-driven guide helps cyclists stay ahead of bicycle thieves.</div>

<div class="description">
Each year, countless bikes disappear from Toronto’s streets, leaving cyclists frustrated and vulnerable. Our analysis digs into patterns and insights across the city, highlighting where and when bicycles are most at risk. With data spanning over a decade, we reveal theft hotspots and provide tips to help keep your bike safe.
</div>

<div class="author-date">Yingke He, Ziheng Zhong<br>November 12, 2024</div>

@@

# you are welcome to use either the CSV file or the GeoJSON file
# GeoJSON format maybe useful if you would like to combine
# an external geospatial dataset
bike_thefts_csv <- read_csv(here("data/clean_data.csv")) |>
  mutate(
    occurence_date = as_date(
      OCC_DATE, format = "%a, %d %b %Y %H:%M:%S GMT"
    ),
    report_date = as_date(
      REPORT_DATE, format = "%a, %d %b %Y %H:%M:%S GMT"
    )
  )
bike_thefts_sf <- read_sf(here("data/Bicycle_Thefts_Open_Data.geojson")) |>
  mutate(
    occurence_date = as_date(
      OCC_DATE, format = "%a, %d %b %Y %H:%M:%S GMT"
    ),
    report_date = as_date(
      REPORT_DATE, format = "%a, %d %b %Y %H:%M:%S GMT"
    )
  )
```

Imagine locking up your bike on a bustling Toronto street, feeling reassured that a strong lock and some common sense will keep it safe. But what if certain seasons, specific bike types, or even certain neighborhoods increase your risk of theft far more than you realize? For many cyclists, bike theft has moved from a distant possibility to an everyday concern, with data revealing just how pervasive the risk can be in Toronto.

In this analysis, we dig into over a decade’s worth of data from the Toronto Police Service to answer key questions: When are bikes at the greatest risk? What makes a bike more likely to be targeted? And where are the theft hotspots cyclists should avoid? Each section of our article addresses a different angle of bike theft in Toronto, from seasonal and weekly theft patterns to variations by bicycle type and value, to pinpointing the city’s most high-risk locations.

With these insights, we aim to empower cyclists with knowledge that can help them safeguard their bikes and inform Cycle Toronto and local authorities about the areas most in need of intervention. Read on as we trace theft trends over the years, reveal the high-risk locations, and provide practical advice to help protect your bike.

@@

```{r interactive_weekly_plot, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=9.5, fig.height=4}
library(plotly)
library(readr)
library(dplyr)
library(here)

# Add layout with the slider and set background colors
fig <- fig %>%
  layout(
    xaxis = list(title = ""),
    yaxis = list(title = "Number of Thefts"),
    sliders = list(
      list(
        active = 0,
        currentvalue = list(prefix = "Day: "),
        steps = steps
      )
    ),
    plot_bgcolor = "#faecd5",     # Set background color for the plotting area
    paper_bgcolor = "#faecd5"     # Set background color for the entire plot
  )

@@

# Calculate yearly totals for annotation
yearly_totals <- data %>%
  group_by(OCC_YEAR) %>%
  summarise(Found_Total = sum(Found, na.rm = TRUE), Stolen_Total = sum(Stolen, na.rm = TRUE))

# Plot for Bikes Found with scaled heights and yearly total annotations
plot_found <- ggplot(data, aes(x = OCC_MONTH_NUM, y = OCC_YEAR, 
                               height = (Found / max_found) * 4, group = OCC_YEAR)) +
  geom_ridgeline(
    fill = "#05595B", color = "white", alpha = 0.5, scale = 1, linewidth = 0
  ) +
  labs(y = "", x = NULL) +  # Remove x-axis title
  theme_minimal() +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(size = 0.5),
    panel.border = element_blank(),
    plot.title = element_text(size = 14, color = "#05595B", hjust = -0.1, vjust = 1.5),  # Smaller title
    panel.background = element_rect(fill = "#faecd5", color = NA),  # Light blue background for plot area
    plot.background = element_rect(fill = "#faecd5", color = NA)    # Light blue background for overall plot
  ) +
  scale_x_discrete(labels = month.abb) +
  ggtitle("Bikes Found") +  # Add title for Bikes Found
  geom_text(data = yearly_totals, aes(x = 12.5, y = OCC_YEAR, label = Found_Total), 
            color = "#05595B", hjust = 1, size = 3, inherit.aes = FALSE)

# Plot for Bikes Stolen with scaled heights, no borders, and yearly total annotations
plot_stolen <- ggplot(data, aes(x = OCC_MONTH_NUM, y = OCC_YEAR, 
                                height = (Stolen / max_stolen) * 4, group = OCC_YEAR)) +
  geom_ridgeline(
    fill = "#603601", color = "white", alpha = 0.5, scale = 1, linewidth = 0
  ) +
  labs(y = "", x = NULL) +  # Remove x-axis title
  theme_minimal() +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(size = 0.5),
    panel.border = element_blank(),
    panel.background = element_rect(fill = "#faecd5", color = NA),  # Light blue background for plot area
    plot.background = element_rect(fill = "#faecd5", color = NA),   # Light blue background for overall plot
    plot.title = element_text(size = 14, color = "#603601", hjust = -0.1, vjust = 1.5)  # Smaller title
  ) +
  scale_x_discrete(labels = month.abb) +
  ggtitle("Bikes Stolen") +  # Add title for Bikes Stolen
  geom_text(data = yearly_totals, aes(x = 12.5, y = OCC_YEAR, label = Stolen_Total), 
            color = "#603601", hjust = 1, size = 3, inherit.aes = FALSE)

@@

@@

```{r animated_bar_plot, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=7, fig.height=4.5}
# Libraries
library(ggplot2)
library(gganimate)
library(readr)
library(here)
library(dplyr)

@@

  # Set background colors for panel and plot
  theme(
    panel.background = element_rect(fill = "#faecd5", color = NA),
    plot.background = element_rect(fill = "#faecd5", color = NA)
  ) +

@@

# Create the animated plot
plot <- data %>%
  ggplot(aes(x = OCC_MONTH_NUM, y = Number_of_Thefts, group = as.factor(OCC_YEAR), color = as.factor(OCC_YEAR))) +
  geom_line() +
  geom_point() +
  scale_color_viridis_d() +
  ylab("Number of Bikes Stolen") +
  xlab("Month") +
  scale_x_continuous(breaks = 1:12, labels = c("January", "February", "March", "April", "May", "June", 
                                               "July", "August", "September", "October", "November", "December")) +
  theme_ipsum() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.background = element_rect(fill = "#faecd5", color = NA),
    plot.background = element_rect(fill = "#faecd5", color = NA)
  ) +
  transition_reveal(OCC_MONTH_NUM) +
  labs(color = "Year")

@@

Map
```{r, include = TRUE, warning = FALSE, message = FALSE}
# Load cleaned coordinates data
coordinates_data <- read.csv(here("data/cleaned_coordinates_data.csv"))

# Plot the map with bike theft locations data
ggplot() +
  geom_sf(data = streets, color = "gray", size = 0.3, alpha = 0.6) + # OSM street map layer
  geom_point(
    data = coordinates_data, aes(x = longitude, y = latitude),
    size = 0.5, color = "red"
  ) + # Bike theft points
  labs(
    title = "Map of Bicycle Theft Locations in Toronto",
    x = "Longitude", y = "Latitude"
  ) +
  coord_sf(
    xlim = c(-79.6303856025883, -79.1180336515019),
    ylim = c(43.5822069220627, 43.8554571861712), expand = FALSE
  ) +
  theme_minimal() +
  theme(
    panel.background = element_blank(), # Remove panel background
    panel.grid.major = element_blank(), # Remove major grid lines
    panel.grid.minor = element_blank(), # Remove minor grid lines
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
